# Reproducible Research: Peer Assessment 1
========================================================

```{r setOptions, echo = FALSE}
opts_chuck$set(echo=TRUE, results = "asis")
```


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv", na.strings = "NA")
head(activity)
dates <- activity$date
dates <- as.Date(as.character(dates), "%Y-%m-%d")
str(dates)
activity$date <- dates
```

## What is mean total number of steps taken per day?
```{r}
sum_steps <- sapply(split(activity, activity$date), function(d){sum(d$steps)})
hist(sum_steps, main="histogram of total steps per day with missing values", xlab = "steps")

mean <- mean(sum_steps,na.rm=TRUE)
median <- median(sum_steps, na.rm=TRUE)
```
Mean total number of steps taken per day is `{r mean}`
Median total number of steps taken per day is `{r median}`

## What is the average daily activity pattern?
```{r}
library(ggplot2)
qplot(interval, steps, data=activity, stat="summary", fun.y="mean",  geom = "line")
interval_steps <- sapply(split(activity, activity$interval), function(d){mean(d$steps, na.rm = TRUE)})
mstep <- which.max(interval_steps)
```
Interval No.`{r names(mstep)}` has the maximum value.

## Imputing missing values
```{r}
sum(is.na(activity$steps))
activity_fill <-data.frame(activity)
for(i in 1:dim(activity)[1]){
    if(is.na(activity$steps[i])){
        activity_fill$steps[i] <- interval_steps[as.character(activity$interval[i])]
    }
}
sum(is.na(activity_fill$steps))

sum_steps_fill <- sapply(split(activity_fill, activity_fill$date), function(d){sum(d$steps)})
hist(sum_steps_fill, main="histogram of total steps per day after filling missing values", xlab = "steps")
```
The mean and median of the total daily number of steps are:
```{r}
mean(sum_steps_fill)
median(sum_steps_fill)
```

These values do differ from the estimates from the first part of the assignment. Imputing missing data decreases the estimates of the total daily number of steps slightly.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
Sys.setlocale(locale = "C")
weekday <- weekdays(activity_fill$date)
for(i in 1:length(weekday)[1]){
    if(weekday[i] %in% c('Sunday','Saturday')){
        weekday[i] <- TRUE
    }
    else
        weekday[i] <- FALSE
}
str(weekday)

activity_fill$weekday<-factor(weekday, labels = c("weekday", "weekend"))
qplot(interval, steps, data=activity_fill, stat="summary", fun.y="mean",  geom = "line", facets = .~weekday, ylab="Number of steps")
```
